---
title: "R Notebook"
output:
  pdf_document: default
  html_document: default
---
##Data import
First we will import the data from the file location
```{r}
MedApp <- read.csv("/Users/harishbharatham/R-Language/Summer2018/BigDataAnalytics/MedAppointmentNoShow/Data/MedAppointmentNoShow.csv")
attach(MedApp)
dim(MedApp)
nrow(MedApp)
summary(MedApp)

```
##Data cleaning
```{r}
# To delete the patientID column 
MedApp$PatientId<- NULL
#changing Gender column into dummy variable female = 1 , male =0
MedApp$Gender<-ifelse(MedApp$Gender=="F",1,0)
#changing No.show column into dummies with yes=1 no=0
MedApp$No.show<-ifelse(MedApp$No.show=="Yes",1,0)
# To seperate date from time
MedApp$Scheduledtime <- substr(ScheduledDay, 11, 20)
MedApp$ScheduledDay <- substr(ScheduledDay, 0, 10)
MedApp$Appointment_Time <- substr(AppointmentDay, 11, 20)
MedApp$Appointment_Date <- substr(AppointmentDay, 0, 10)
View(MedApp)

#To mention categorical columns as characters 
cols_to_change = c(2, 7:11,13)
for(i in cols_to_change){
     class(MedApp[,i]) = "character"
}
#To mention categorical columns as characters 
View(MedApp)
```


####Creating 'DaysWaited' feature
We need to quantify the Appointment date and Scheduled date. So we create a variable that will substract both dates and provides us a waiting time in days which we can use in the model for prediction.
```{r}
MedApp.scheduled <- as.Date(MedApp$ScheduledDay)
MedApp.Appmtmnt <- as.Date(MedApp$AppointmentDay)
DayWaited <- as.integer(c(MedApp.Appmtmnt - MedApp.scheduled))
MedApp$DayWaited <- DayWaited

columnName <- c("Gender","DayWaited","Age","Scholarship","Hipertension",
                "Diabetes","Alcoholism","Handcap","SMS_received","No.show")
MedApp <- MedApp[columnName]

cols_to_change = c(1:10)
for(i in cols_to_change){
     class(MedApp[,i]) = "integer"
}
View(MedApp)

str(MedApp)

```

##Data Split
We need to create test and train data for the model. We create a sample size and we keep the seed 123 as random, but to get the same results we keep the 123 constant.
```{r}
sampleSize <- floor(0.7 * nrow(MedApp))
set.seed(123)
index <- sample(seq_len(nrow(MedApp)), size = sampleSize)

train <- MedApp[index, ]
test <- MedApp[-index, ]
```

##Logistic Regression
We have the test and train split, we will create the logistic regression model and run it on the train split of the data.
```{r}
b <- step(glm(No.show~1,data=train,family='binomial'),
          scope=list(lower=~1,upper=~Gender+DayWaited+Age+Scholarship+Hipertension
                     +Diabetes+Alcoholism+Handcap+SMS_received),
          direction='both')

summary(b)
```

Now we will run the model with the test data to generate the prediction
```{r}
b.prediction <- predict(b,newdata=test,type='response')
b.prediction[b.prediction >= 0.4] <- 1
b.prediction[b.prediction < 0.4] <- 0
b.prediction <- as.vector(b.prediction)
```

Now let's create a confusion matrix to easily understand the model performance. For that, we need to import a library, caret.
```{r}
library(caret)
library(e1071)
b_conf <- confusionMatrix(b.prediction,test$No.show)
b_conf
```
##Neural Networks
Now let's create and run the Artificial Neural Networks model and run it on train and test data.
```{r}
library(nnet)
set.seed(1234)
nn.p <-nnet(No.show~., data= train, size = 4, rang= .1,decay=.0005,maxit=200)
nn.prediction<-predict(nn.p,test)
nn.prediction[nn.prediction<0.4] <-0
nn.prediction[nn.prediction>=0.4] <-1
```

Let's create a confusion matrix to evaluate the performance of the model.
```{r}
nn_conf <- confusionMatrix(nn.prediction,test$No.show)
nn_conf

```

##Results
As we can see, the ANN performed better than logistic, but both the results are close.
```{r}



accuracy <- nn_conf$overall['Accuracy']
precision <- nn_conf$byClass['Pos Pred Value']
recall <- nn_conf$byClass['Sensitivity']
F1 <- 2*((precision*recall)/(precision+recall))
r <- data.frame(rbind(accuracy,precision,recall,F1))


accuracy <- b_conf$overall['Accuracy']
precision <- b_conf$byClass['Pos Pred Value']
recall <- b_conf$byClass['Sensitivity']
F1 <- 2*((precision*recall)/(precision+recall))
r<- cbind(r,rbind(accuracy,precision,recall,F1))
names(r) <- c("ANN Model","Logistic Model")

r
```

##Sample data
Created Sample dataframes to test the model
```{r}
x <- data.frame("Gender" = 0,"DayWaited"= 10,"Age" = 25,"Scholarship" = 0,"Hipertension" = 0,
                "Diabetes" = 0,"Alcoholism"= 1,"Handcap" = 0,"SMS_received" = 1)

nn.prediction<-predict(nn.p,x)
nn.prediction
nn.prediction[nn.prediction<0.4] <-"Show"
nn.prediction[nn.prediction>=0.4 && nn.prediction!="Show"] <-"NoShow"
nn.prediction
```

```{r}
y <- data.frame("Gender" = 1,"DayWaited"= 20,"Age" = 22,"Scholarship" = 0,"Hipertension" = 0,
                "Diabetes" = 0,"Alcoholism"= 1,"Handcap" = 0,"SMS_received" = 0)
nn.prediction<-predict(nn.p,y)
nn.prediction
nn.prediction[nn.prediction<0.4] <-"Show"
nn.prediction[nn.prediction>=0.4 && nn.prediction!="Show"] <-"NoShow"
nn.prediction

```

